### 2015-09-16 Manuscript/SW_early_life_telomeres_tables_and_figures v9.Rmd
Discussion and results work

**Figure 2** Temporal variation in early-life telomere dynamics in the Seychelles warbler based on cross-sectional **(A,B)** and longitudinal **(C,D)** data (see Fig. 1 and main text for details). Points and error bars represent mean and standard error telomere length for all birds born in each summer (blue) or winter (red) breeding season. In **A** and **C** points are ordered temporally, while in **B** and **D** telomere length/loss for each season is plotted against island-wide food availability. Lines and shaded areas represent fitted values and 95% confidence limits from a linear regression.

@@

**Figure 5** Temporal variation in early-life telomere dynamics and lifespan in the Seychelles warbler, based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent means and $\pm$ s.e., respectively.

@@

```{r Figure 1,fig.width=6,fig.height=9}
######################
#1A
######################

@@

Fig1A <- ggplot(ddFig1A,
       aes(x = Ageclass,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.5,y=6.2)

@@

Fig1B <- ggplot(ddFig1B,
       aes(x = Ageclass,
           y = TROCmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.5,y=0.0018)

@@

```{r Figure 2,fig.width=9,fig.height=9}

Fig2A <- ggplot(juvseason,
       aes(x = CatchYear,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse),
                  position = position_dodge(width = 0.2)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'A',x=1995,y=10)

@@

Fig2B <- ggplot(juvseason,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se=T,col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'B',x=-1.5,y=8)

@@

Fig2C <- ggplot(temp,
       aes(x = CatchYear,
           y = TROCmean,
           col = Season)) +
  ylab('Corrected telomere loss') + 
  xlab('Year') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse),
                  position = position_dodge(width = 0.2)) +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=1995,y=0.005)

####################################
Fig2D <- ggplot(temp,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm', se=T, col = 'darkgrey') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'D',x=-1.5,y=0.0075)

multiplot(Fig2A,Fig2B,Fig2C,Fig2D,layout=matrix(1:4,2,2))

@@

Fig3C <- ggplot(chicks,
                aes(x = Tarsus,
                    y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point(size=3) +
  stat_smooth(method = 'lm',col = 'black') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',label = 'C',x=19,y=15)

@@

Fig5A <- ggplot(juvseason,
                aes(x = TLKBmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TLKBmean-TLKBse,xmax = TLKBmean+TLKBse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Telomere length (kb)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = wes_palette("Royal1")) +
  annotate('text',x=2.5,y=12.5,label='A')

Fig5B <- ggplot(loss.season,
                aes(x = TROCmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TROCmean-TROCse,xmax = TROCmean+TROCse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Corrected telomere loss') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 11), 
        axis.text.y = element_text(size = 11), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = wes_palette("Royal1")) +
  stat_smooth(method = 'lm',col = 'darkgrey',fullrange = T) +
  annotate('text',x=-0.004,y=14,label='B')

@@

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig6A <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'FL'))
Fig6B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

@@

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'SA'))
Fig6C <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('darkred','red')) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='C')

>>>>

**Figure 2** Temporal variation in early-life telomere dynamics in the Seychelles warbler based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent mean and standard error telomere length for all birds born in each summer (black) or winter (grey) breeding season, in relation to island-wide food availability. Trend lines represent fitted values from a linear regression.

@@

**Figure 5** Temporal, cohort-level variation in early-life telomere dynamics and lifespan in the Seychelles warbler, based on cross-sectional **(A)** and longitudinal **(B)** data (see Fig. 1 and main text for details). Points and error bars represent means and $\pm$ s.e., respectively, of all birds born in each summer (black) or winter (grey) breeding season.

@@

```{r Figure 1,fig.width=5,fig.height=9}
######################
#1A
######################

@@

Fig1A <- ggplot(ddFig1A,
       aes(x = Ageclass,
           y = TLKBmean)) +
  ylab('Telomere length (kb)') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  annotate('text',label = 'A',x=0.6,y=6.2)

@@

Fig1B <- ggplot(ddFig1B,
       aes(x = Ageclass,
           y = TROCmean)) +
  ylab('Corrected telomere loss') + 
  xlab('') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.1,0.9),
        legend.title = element_blank(),
        legend.text = element_text(size = 15)) +
  geom_pointrange(aes(ymin=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  annotate('text',label = 'B',x=0.6,y=0.0018)

@@

```{r Figure 2,fig.width=5,fig.height=9}

@@

Fig2A <- ggplot(juvseason,
       aes(x = Insect,
           y = TLKBmean,
           col = Season)) +
  ylab('Telomere length') + 
  xlab('Food availability') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  geom_pointrange(aes(ymin=TLKBmean-TLKBse,ymax=TLKBmean+TLKBse)) +
  stat_smooth(method = 'lm',se = F,col = 'black') +
  scale_color_manual(values = c('black',grey(0.4))) +
  annotate('text',label = 'A',x=-1.5,y=8)

@@

####################################
Fig2B <- ggplot(temp,
       aes(x = Insect,
           y = TROCmean,
           col = Season)) +
  geom_pointrange(aes(min=TROCmean-TROCse,ymax=TROCmean+TROCse)) +
  ylab('Corrected telomere loss') + 
  xlab('Food availability') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none') +
  stat_smooth(method = 'lm',se = F,  col = 'black') +
  scale_color_manual(values = c('black',grey(0.4))) +
  annotate('text',label = 'B',x=-1.5,y=0.0075)

multiplot(Fig2A,Fig2B)

@@

Fig3C <- ggplot(chicks,
                aes(x = Tarsus,
                    y = TLKB)) +
  ylab('Telomere length (kb)') + 
  xlab('Tarsus length (mm)') +
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16)) +
  geom_point(size=3) +
  stat_smooth(method = 'lm',se = F,col = 'black') +
  scale_color_manual(values = c('black',grey(0.4))) +
  annotate('text',label = 'C',x=19,y=15)

@@

Fig5A <- ggplot(juvseason,
                aes(x = TLKBmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TLKBmean-TLKBse,xmax = TLKBmean+TLKBse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Telomere length (kb)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 12), 
        axis.text.y = element_text(size = 12), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = c('black',grey(0.4))) +
  annotate('text',x=2.5,y=12.5,label='A')

Fig5B <- ggplot(loss.season,
                aes(x = TROCmean,
                    y = Lifespan,
                    col = Season)) +
  geom_pointrange(aes(ymin=Lifespan-Lifespanse,ymax=Lifespan+Lifespanse)) +
  geom_errorbarh(aes(xmin = TROCmean-TROCse,xmax = TROCmean+TROCse)) +
  theme_classic() +
  ylab('Remaining life') + 
  xlab('Corrected telomere loss') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 11), 
        axis.text.y = element_text(size = 11), 
        axis.title.y = element_text(size = 14,vjust=0.8),
        axis.title.x = element_text(size = 14),
        legend.position = 'none') +
  scale_color_manual(values = c('black',grey(0.4))) +
  stat_smooth(method = 'lm',se = F,col = 'black',fullrange = T) +
  annotate('text',x=-0.004,y=14,label='B')

@@

s1 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = chicks)
Fig6A <- ggsurv(s1,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = c(0.8,0.8),
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.4))) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='A')

###########################################
s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'FL'))
Fig6B <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.4))) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='B')

@@

s2 <- survfit(Surv(time = Lifespan+1, event = Died)~TLF, data = subset(FlSAall,Ageclass == 'SA'))
Fig6C <- ggsurv(s2,plot.cens=F) + 
  theme_classic() +
  ylab('Proportion survived') + 
  xlab('Age (years)') + 
  theme_classic() +
  theme(axis.text.x = element_text(size = 14), 
        axis.text.y = element_text(size = 14), 
        axis.title.y = element_text(size = 16,vjust=0.8),
        axis.title.x = element_text(size = 16),
        legend.position = 'none',
        legend.text = element_text(size = 12),
        legend.title = element_blank()) +
  scale_colour_manual(name = 'TL',
                      values = c('black',grey(0.4))) +
  guides(linetype=F) +
  annotate('text',x=0.5,y=1.1,label='C')
